Centralizers of higher enveloping algebras of Lie algebras

### 2020-09-07 draft.tex
revised structure and added stubs for some easy applications

%Recently, a generalization of the concept of the enveloping algebra of a Lie algebra has emerged: each dg Lie algebra $\fg$ has an enveloping $E_n$ algebra $\UU_n\fg$, for every $n$, so that $\UU_1\fg$ is the usual enveloping algebra $U\fg$. Building on prior work \cite{BD, AF} [Owen's thesis?], Knudsen \cite{Knudsen} developed these ideas extensively and offered a useful factorization algebra model for the enveloping $E_n$ algebra. Our central goal here is to generalize the following results about the enveloping algebra $U\fg$ to $\UU_n\fg$, for all $n$.
%
%\begin{enumerate}
%\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where $\CC^{N}$ denotes the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ is the Weyl group of $SL_{N+1}$, and
%\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$. That is to say $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$.
%\end{enumerate}
%
%Our results pertains to (1) by taking $N \to \infty$, and to (2) through the derived {\em centralizer} of the map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra. Note that the usual derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.
%
%We will review these notions and describe our results in detail Min the next section. When $n=1$,  \owen{state the classical thing here!}
%
%Our primary tool here is the Loday-Quillen-Tsygan (LQT) theorem which describes the large $N$ limit of the Lie algebra homology of $\gl_N(A)$, for a dg associative algebra $A$. Since the enveloping $E_n$ algebra of a Lie algebra can be constructed as a locally constant factorization algebra made out of the shifted Chevalley-Eilenberg chains \cite{}, the relevance of the LQT theorem in this context  should be apparent. From this point of view, one can think of our main result as an $E_n$ generalization of the LQT theorem. We hope these ideas have exciting applications in quantum field theory and random matrix theory, as well as to other large $N$ limit phenomena.

@@

@@

Likewise, on each open set $V \subset \RR^k$, 
we obtain a quasi-isomorphism
\[
\bZ_k(\fg_S)(V) \cong \clies( \Omega^*(V) \otimes \fg_S, \bU_k \fg_S) \xto{\simeq} \clies(\Omega^*(V \times S) \otimes \fg, \bU_n(V \times S))
\]
and this map is functorial in opens.
The righthand side is, however, the pushforward $\int_S \bZ_n \fg$.
\owen{Hmmm, that map goes the wrong way, but I'm pretty sure it provides the inverse-up-to-qiso we need.}
\end{proof}

@@

\section{Universal bulk TFT}
\label{sec: TFT}
%
%Given an $E_{n+1}$-algebra one can speak of its $\infty$-category of left (or right) modules.
%This is obtained by ``forgetting" the $E_{n+1}$-algebra to an $E_1$-algebra then considering the left-modules for this $E_1$-algebra.
%
%The $E_n$-center, or $E_n$-Hochschild cohomology, of an $E_n$-algebra $A$ is the {\em universal} $E_{n+1}$-algebra which acts on $A$. 
%In Section \ref{sec:swiss} below, we will recall an elegant operadic interpretation of this universal picture and interpret it in terms of {\em constructible factorization algebras}. 
%In the case of $E_n$-enveloping algebras, we will provide a strict factorization model of this universal situation in terms of a constructible factorization algebra on an $(n+1)$-dimensional 
%manifold with boundary. 
%
%We will furthermore provide an interpretation of this constructible factorization algebra in terms of the quantization of a certain topological field theory. 
%The key insight is that there is a {\em classical limit} of the $E_n$-center $\fZ_n(\fg)$ which is equivalent to the observables of an $(n+1)$-dimensional topological field theory known as {\em BF theory}. 
%Likewise, there is a classical limit of $\UU_n(\fg)$ which can be viewed as a boundary condition for BF theory.
%
%In this sense, topological BF theory is the universal $(n+1)$-dimensional field theory which has the $E_n$-enveloping algebra as a boundary condition. 
%This partially confirms the terminology used in \cite{ButsonYoo}  of topological BF theory as the ``universal bulk theory" of $\UU_n(\fg)$.

@@

\begin{prp}
Filtered swiss cheese.
Gr is a P_0 swiss cheese algebra.
\end{prp}

@@

\section{Examples?}

\greg{Add the Koszul duality example. Add examples coming from topological spaces and abelian Lie algebras}

\subsection{$E_n$-calculus}
\bibliographystyle{alpha}  
\bibliography{centralize}

>>>>

\section{Introduction}

Recently, a generalization of the concept of the enveloping algebra of a Lie algebra has emerged: 
each dg Lie algebra $\fg$ has an enveloping $E_n$ algebra $\UU_n\fg$, for every $n$, 
so that $\UU_1\fg$ is the usual enveloping algebra $U\fg$. 
Building on prior work \cite{BD, AF} [Owen's thesis?], 
Knudsen \cite{Knudsen} developed these ideas extensively and offered a useful factorization algebra model for the enveloping $E_n$ algebra. 
Our central goal here is to generalize the some well-known results about the enveloping algebra $U\fg$ to $\UU_n\fg$, for all $n$, such as
\begin{enumerate}
\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where $\CC^{N}$ denotes the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ is the Weyl group of $SL_{N+1}$, and
\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$. That is to say $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$.
\end{enumerate}
\owen{And keep going}
%
%Our results pertains to (1) by taking $N \to \infty$, and to (2) through the derived {\em centralizer} of the map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra. Note that the usual derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.
%
%We will review these notions and describe our results in detail Min the next section. When $n=1$,  \owen{state the classical thing here!}
%
%Our primary tool here is the Loday-Quillen-Tsygan (LQT) theorem which describes the large $N$ limit of the Lie algebra homology of $\gl_N(A)$, for a dg associative algebra $A$. Since the enveloping $E_n$ algebra of a Lie algebra can be constructed as a locally constant factorization algebra made out of the shifted Chevalley-Eilenberg chains \cite{}, the relevance of the LQT theorem in this context  should be apparent. From this point of view, one can think of our main result as an $E_n$ generalization of the LQT theorem. We hope these ideas have exciting applications in quantum field theory and random matrix theory, as well as to other large $N$ limit phenomena.

\subsection{Overview of the paper}

\subsection{Acknowledgments}

\subsection{Notations}

@@

\section{A few examples of the central result}

\subsection{Abelian Lie algebras and an $E_n$ HKR theorem}

Given a free dg commutative algebra $A = \Sym(V^\bullet)$ generated by a cochain complex $V^\bullet$,
one can view it as an $E_n$ algebra via the forgetful functor determined by the operad map $E_n \to E_\infty$.
We will blithely view any commutative algebra as an $E_n$ algebra for any $n$ via this forgetting.

We also know that for an abelian dg Lie algebra~${\frak a}$, 
\[
\cliels({\frak a}) = \Sym({\frak a}[1])
\]
and hence provides a free dg commutative algebra.
This feature can be leveraged to prove the following result about $E_n$-centers of free commutative algebras.

\begin{prp}
Given a free dg commutative algebra $A = \Sym(V^\bullet)$ generated by a cochain complex $V^\bullet$,
its center ${\frak Z}_n(A)$ is equivalent to the $E_{n+1}$ algebra of its shifted polyvector fields 
\[
{\rm Poly}_n(A) = \Sym(V^\bullet \oplus (V\bullet)^\vee[-n]).
\]
\end{prp}

\begin{proof}
A cochain complex $V^\bullet$ is an abelian dg Lie algebra;
indeed, so is any shift $V^\bullet[k]$. 
Let $\fg_V$ denote $V^\bullet[-n+1]$ viewed as an abelian dg Lie algebra.
The Poincar\'e lemma for compactly supported de Rham forms on $\RR^n$ provides a canonical quasi-isomorphism
\[
\Omega^*_c(\RR^n) \xto{\simeq} \RR[-n]
\]
by the natural quotient map and hence determines a canonical quasi-isomorphism
\[
\cliels(\Omega^*_c(\RR^n) \otimes \fg_V) \xto{\simeq} \Sym(V^\bullet).
\]
This map makes sense on any ball in $\RR^n$ and thus implies that $\bU_n \fg_V$ is quasi-isomorphic as a $\Disk_n$ algebra to~$\Sym(V^\bullet)$.
Applying our central theorem, we see
\begin{align*}
\bZ_n(\fg_V) &= \clies(\Omega^* \otimes \fg_V, \bU_n \fg_V) \\
&= \Sym((\Omega^* \otimes \fg_V)^\vee[-1]) \otimes \Sym(\Omega^*_c \otimes \fg_V[1])\\
& \simeq \Sym(V^\bullet \oplus (V\bullet)^\vee[-n])
\end{align*}
because
\[
\Omega^*(\RR^n) \otimes \fg_V \simeq V[-n+1]
\]
so
\[
\Sym( (\Omega^*(\RR^n) \otimes \fg_V)^\vee[-1]) \simeq \Sym( V^\vee[-n]).
\]
This identification concludes the argument.
\end{proof}

This result is the simplest case of a generalized Hochschild-Kostant-Rosenberg theorem for commutative algebras viewed as $E_n$ algebras.
(The classic theorem proves the $n=1$ situation.)
Such a result is proved by very different methods in~\cite{CalWil}.
Now we swiftly describe how to use methods like those in this paper to prove a version of this powerful result.

Our approach applies only to a certain class of dg commutative algebras,
due to a technical hypothesis necessary to apply our methods.
We need as input an augmented dg commutative algebra $\epsilon: A \to k$ that admits a semifree resolution of the form $(\Sym(V^\vee), \partial)$, where $V$ is some graded vector space.
In other words, we want to know that the generators can be realized as the graded linear dual of some other graded vector space.
We say such an $A$ is of {\em type~($\star$)}.

\begin{prp}
Let $A$ be an augmented dg commutative algebra of type~($\star$).
Then 
\[
{\frak Z}_n(A) \simeq {\rm Poly}_n(A) \simeq \Sym_A(\TT_A[-n]).
\]
\end{prp}

\owen{This proposition should be phrased correctly. The RHS isn't quite correct unless $A$ is good enough.}

\begin{proof}
\owen{The method described in the email.}
\end{proof}

\owen{Also prove that the $E_n$ center of $\clies(\fg)$ is also $\clies(\fg, \UU_n \fg)$}

\begin{cor}
\label{cor: koszul dual to main theorem}
For any $L_\infty$ algebra $\fg$,
there is a quasi-isomorphism
\[
{\frak Z}_n(\clies(\fg)) \simeq \clies(\fg, \UU_n \fg).
\]
\end{cor}

\subsection{Reductive Lie algebras and a generalized Harish-Chandra homomorphism}

\begin{prp}
For $\fg$ a reductive \owen{semisimple?} Lie algebra with Cartan subalgebra~$\fh$ and Weyl group~$W$, 
there is a quasi-isomorphism of $E_{n+1}$ algebra
\[
{\frak Z}_n(\UU_n \fg) \simeq H^*_{\rm Lie}(\fg) \otimes \Sym(\fh[1-n])^W.
\]
\end{prp}

\begin{proof}
\owen{In brief
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies(\fg, \Sym(\fg[1-n])
\]
and then apply theorems about Lie algebra cohomology plus Chevalley restriction theorem.}
\end{proof}

\subsection{The large $N$ limit of the center ${\frak Z}_n(\UU_n {\frak gl}_N)$ and cyclic cohomology}

\owen{Point out how to use LQT map and deduce existence of $E_n$ algebra structure on a shifted copy of cyclic cohomology of~$\Omega^*_c$}

\subsection{Free Lie algebras and free $E_n$ algebras}

\owen{Describe what we can of the center of a free $E_n$ algebra thanks to Ben's theorem}

@@

Likewise, on each open set $V \subset \RR^k$, 
we obtain a quasi-isomorphism
\[
\bZ_k(\fg_S)(V) \cong \clies( \Omega^*(V) \otimes \fg_S, \bU_k \fg_S) \xto{\simeq} \clies(\Omega^*(V \times S) \otimes \fg, \bU_n(V \times S))
\]
and this map is functorial in opens.
The righthand side is, however, the pushforward $\int_S \bZ_n \fg$.
\owen{That map is not the canonical map as it goes the ``wrong way,'' but I'm pretty sure it provides the inverse-up-to-qiso we need.}
\end{proof}

@@

\subsection{Koszul duality for enveloping $E_n$ algebras}

\owen{Use corollary~\ref{cor: koszul dual to main theorem} to exhibit a "dual" Swiss cheese pair and then compute the relative tensor product over the center, which will be the unit $E_n$ algebra. Draw a picture of the strip.}

\section{Centers and topological field theories}
\label{sec: TFT}
%
%Given an $E_{n+1}$-algebra one can speak of its $\infty$-category of left (or right) modules.
%This is obtained by ``forgetting" the $E_{n+1}$-algebra to an $E_1$-algebra then considering the left-modules for this $E_1$-algebra.
%
%The $E_n$-center, or $E_n$-Hochschild cohomology, of an $E_n$-algebra $A$ is the {\em universal} $E_{n+1}$-algebra which acts on $A$. 
%In Section \ref{sec:swiss} below, we will recall an elegant operadic interpretation of this universal picture and interpret it in terms of {\em constructible factorization algebras}. 
%In the case of $E_n$-enveloping algebras, we will provide a strict factorization model of this universal situation in terms of a constructible factorization algebra on an $(n+1)$-dimensional 
%manifold with boundary. 
%
%We will furthermore provide an interpretation of this constructible factorization algebra in terms of the quantization of a certain topological field theory. 
%The key insight is that there is a {\em classical limit} of the $E_n$-center $\fZ_n(\fg)$ which is equivalent to the observables of an $(n+1)$-dimensional topological field theory known as {\em BF theory}. 
%Likewise, there is a classical limit of $\UU_n(\fg)$ which can be viewed as a boundary condition for BF theory.
%
%In this sense, topological BF theory is the universal $(n+1)$-dimensional field theory which has the $E_n$-enveloping algebra as a boundary condition. 
%This partially confirms the terminology used in \cite{ButsonYoo}  of topological BF theory as the ``universal bulk theory" of $\UU_n(\fg)$.

@@

\begin{prp}
Filtered swiss cheese.
$Gr$ is a $P_0$ swiss cheese algebra.
\end{prp}

@@

\bibliographystyle{alpha}  
\bibliography{centralize}
